---
title: "Plot Barcharts from Is there a gender bias in Chemical Sciences publishing?"
output:
  
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

by RSC Data Science

##Is there a gender bias in Chemical Sciences publishing?
###R Notebook showing example code

```{r echo=FALSE,include=FALSE,warning=FALSE,message=FALSE, cache=FALSE}
library(ggplot2)
library(data.table)
library(forcats)
library(reshape2)
library(plyr)
library(knitr)
library(weights)

library(effects)
library(vcd)
library(nnet)
library(actuar)

source('./genderDiversitySharedFunctions.R')
source('./genderAssignmentByName.R')

```

###Example 1: Running gender assignment code

Code is in ./genderAssignmentByName.R
```{r echo=FALSE,warning=FALSE,message=FALSE}
inputdf = read.csv("data/fileWithNameInIt.csv")
inferredGender<-assignGenderToVector( inputdf$firstname)
inputdf$gender <-inferredGender$GENDER
kable(inputdf)

```

###Example 2: Basic bar chart with significance bars, and baseline
```{r echo=FALSE,warning=FALSE,message=FALSE}

inputdf = read.csv("data/s1_authorrevieweraverages.csv", fileEncoding="UTF-8-BOM")

#important that this is set correctly
baseFemalePercentage <-getvaluefromdf("s1_baselines","Level","All authors","FemalePercentage")

inputdf$Level <- factor(inputdf$Level, levels = inputdf[,"Level"])
inputdf<-addTotalAndPercentageColumnsToDF(inputdf)
inputdf<-calculateBinomialsDF(inputdf,baseFemalePercentage)

baseFemaleProportion <-(baseFemalePercentage/100)


totalplotdf<-maketotalplotdf(inputdf)


plot<-plotTotalGraph(totalplotdf, baseFemalePercentage,"Number of original submissions","Total number")
plot + coord_flip()

percentplotdf<-makepercentplotdf(inputdf)
plot_1_2b <-plotPercentageGraph(percentplotdf, baseFemalePercentage,"Number of original submissions","Percentage", "Corresponding Authors")
plot_1_2b + coord_flip()

displaytable(inputdf,"Number of original submissions")

```

###Example 3: Running code for determining if difference between weighted averages is significant


```{r echo=FALSE,warning=FALSE,message=FALSE}

inputdf = read.csv("data/s2_CorrespondingAuthorVsNoSubmissions.csv", fileEncoding="UTF-8-BOM")

#Make list of each Number of Submissions repeated by the number of females to make mean, var and t-test calculations more straightforward without other weighting needed
expandedfemales <- rep(inputdf$Level, inputdf$Female)
# and for males
expandedmales <- rep(inputdf$Level, inputdf$Male)

femaleMean <- mean(expandedfemales)
maleMean <- mean(expandedmales)
n_female<-length(expandedfemales)
n_male<-length(expandedmales)
femaleStandardError<- sd(expandedfemales)/sqrt(n_female)
maleStandardError<- sd(expandedmales)/sqrt(n_male)
ttestresults<-t.test(expandedfemales,expandedmales)
pvalue<-ttestresults$p.value

significance<-"not significant"
if (pvalue< 0.05) {
  significance<-"significant"
}

```
mean and sd functions of R base library can be used to calculated means and standard errors for females and males e.g. Numbers of submissions from file data/s2_CorrespondingAuthorVsNoSubmissions.csv, and the t.test function to calculate the p-value to determine if they are significantly different:

* mean for females = `r toString(round(femaleMean,2))`, standard error = `r toString(round(femaleStandardError,3))`, n = `r toString(n_female)`
* mean for males = `r toString(round(maleMean,2))`, standard error = `r toString(round(maleStandardError,3))`, n = `r toString(n_male)`
* this difference is `r significance` (two sample t-test, p = `r toString(pvalue)`).

###Example 4: Binomial GLM model with one variable

Here is an example of how we ran a binomial glm model with one variable.

For this question we have filtered out:

* instances where the cited corresponding gender is Unknown.
* instances where the citing corresponding gender is Unknown.
* instances where the citing corresponding author matches the cited corresponding or first authors

```{r echo=FALSE,warning=FALSE,message=FALSE}

OutcomeName<- "cited.correspondingauthorgender"
Variable1Name<-"citing.correspondingauthorgender"
rawdataDF = read.csv("data/RSCCitedToCitingFile.csv")

inputDF<-rawdataDF[which(rawdataDF$cited.correspondingauthorgender!="U"),]
inputDF<-inputDF[which(inputDF$citing.correspondingauthorgender!="U"),]
inputDF<-inputDF[which(inputDF$SelfCitingCorrespondingAuthor!="Self-cited"),]

inputDF$cited.correspondingauthorgender <- factor(inputDF$cited.correspondingauthorgender, levels = c("M","F"))

inputDF$citing.correspondingauthorgender <- factor(inputDF$citing.correspondingauthorgender, levels = c("M","F"))

glm(cited.correspondingauthorgender~citing.correspondingauthorgender,data=inputDF,family = "binomial")->modelToUse1
summaryoutput1<-summary(modelToUse1)
as.data.table(effect("citing.correspondingauthorgender",data=inputDF,mod=modelToUse1,confidence.level=.8))->effectTable1
#head(effectTable)

anovasummary<-anova(modelToUse1,test="Chisq")
PValueForModel<-anovasummary[2,"Pr(>Chi)"]
SignificanceForModel<-GetSignificanceFromPValue(PValueForModel)

Variable1plot<- ggplot(effectTable1,aes(y=fit,ymin=lower,ymax=upper,x=citing.correspondingauthorgender,group=1))+geom_point()+geom_line()+geom_errorbar(alpha=0.7,width=0.1)+ylab("Proportion of female cited corresponding authors")+xlab("Gender of Citing Corresponding Author")+ggtitle(paste0("Effects on Proportion of Female Citing Corresponding Authors of Cited Corresponding Author Gender"))+theme_bw()

Variable1plot
```

Figure - Binomial GLM model of cited corresponding author gender of citations and citing corresponding author gender (Model: `r OutcomeName` ~ `r Variable1Name`). ANOVA: `r Variable1Name` is `r toString(SignificanceForModel)` (p = `r toString(PValueForModel)`)

```{r echo=FALSE,warning=FALSE,message=FALSE}

resultsDF<- addPopulationSizeToDataFrame(effectTable1,modelToUse1,"citing.correspondingauthorgender","")

resultsDF <-addPValuesAndSignificanceOneVariable(resultsDF,t(unname(coef(summaryoutput1)[,4])))


kable(resultsDF)

```

* From the plot and the figures it can be seen that the Gender of the Citing Corresponding Author does have significant effect on the proportion of female cited corresponding authors - the value for female citing corresponding authors is **`r toString(round(effectTable1[1,"fit"],3))`** and that for male citing corresponding authors is lower at **`r toString(round(effectTable1[2,"fit"],3))`**. 


###Example 5: Binomial GLM model with two variables

Here is an example of how we ran a glm model with one variable.

For this question we have filtered out:

* instances where the cited first author gender is Unknown.
* instances where the citing corresponding author gender is Unknown.
* instances where the citing first author is the same as the citing corresponding author
* instances where the citing corresponding author matches the cited corresponding or first authors
* instances where the citing first author matches the cited corresponding or first authors

```{r echo=FALSE,warning=FALSE,message=FALSE}

Variable1Name<-"citing.correspondingauthorgender"
Variable2Name<-"citing.firstauthorgender"

inputDF <- rawdataDF[which(rawdataDF$cited.correspondingauthorgender!="U"),]

inputDF <- inputDF[which(inputDF$SameCitingCorrespondingAndFirstAuthor=="False"),]

inputDF <- inputDF[which(inputDF$citing.correspondingauthorgender!="U"),]
inputDF <- inputDF[which(inputDF$citing.firstauthorgender!="U"),]
inputDF<-inputDF[which(inputDF$SelfCitingCorrespondingAuthor!="Self-cited"),]
inputDF<-inputDF[which(inputDF$SelfCitingFirstAuthor!="Self-cited"),]

inputDF$cited.correspondingauthorgender <- factor(inputDF$cited.correspondingauthorgender, levels = c("M","F"))

inputDF$citing.correspondingauthorgender <- factor(inputDF$citing.correspondingauthorgender, levels = c("M","F"))

inputDF$citing.firstauthorgender <- factor(inputDF$citing.firstauthorgender, levels = c("M","F"))

glm(cited.correspondingauthorgender~citing.correspondingauthorgender*citing.firstauthorgender,data=inputDF,family = "binomial")->modelToUse1_2
summaryoutput1_2<-summary(modelToUse1_2)

anovasummary<-anova(modelToUse1_2,test="Chisq")
PValueForModel<-anovasummary[4,"Pr(>Chi)"]
SignificanceForModel<-GetSignificanceFromPValue(PValueForModel)

as.data.table(effect("citing.correspondingauthorgender*citing.firstauthorgender",data=inputDF,mod=modelToUse1_2,confidence.level=.8))->effectTable1_2

resultsDF<- addPopulationSizeToDataFrame(effectTable1_2,modelToUse1_2,Variable1Name,Variable2Name)

resultsDF<-addPValuesAndSignificanceTwoVariables(resultsDF,coef(summaryoutput1_2)[,4],Variable1Name,Variable2Name,levels(resultsDF[[Variable1Name]])[1],levels(resultsDF[[Variable2Name]])[1]) 

ggplot(resultsDF,aes(y=fit,ymin=lower,ymax=upper,x=citing.correspondingauthorgender,group=1))+
  geom_point()+
  geom_line()+
  geom_errorbar(alpha=0.7,width=0.1)+
  theme(legend.position="bottom",legend.direction="horizontal")+
  ylab("Proportion of female cited corresponding authors")+
  xlab("Gender of Citing Corresponding Author")+
  ggtitle(paste0("Breakdown by citing first author gender in separate plots"))+
  facet_wrap(~citing.firstauthorgender)+theme_bw()

```

Figure - Binomial GLM model of cited corresponding author gender of citations, citing corresponding author gender and citing first author gender (Model: `r OutcomeName` ~ `r Variable1Name` \* `r Variable2Name`). ANOVA: `r Variable1Name` is `r toString((GetSignificanceFromPValue(anovasummary[2,"Pr(>Chi)"])))` (p = `r toString(anovasummary[2,"Pr(>Chi)"])`); `r Variable2Name` is `r toString(GetSignificanceFromPValue(anovasummary[3,"Pr(>Chi)"]))` (p = `r toString(anovasummary[3,"Pr(>Chi)"])`); `r Variable1Name` \* `r Variable2Name` is `r toString(SignificanceForModel)` (p = `r toString(PValueForModel)`)

```{r echo=FALSE,warning=FALSE,message=FALSE}

kable(resultsDF)

mosaic(~ citing.correspondingauthorgender+citing.firstauthorgender, data = inputDF, labeling = labeling_values,rot_labels=c(0,90,0,0))
```

###Example 6: Multinomial GLM model with two variables

Here is an example of how we ran a multinomial glm model with two variables.

For these questions we have filtered out:

* instances where the corresponding author gender is Unknown.
* instances where the reviewer gender is Unknown.
* instances where the minor decision is "recommend after revision", "recommended after revision" or "revisions required"
    
```{r echo=FALSE,warning=FALSE,message=FALSE}

rawdataDF <- read.csv("data/s1_OneRowPerReview.csv")
rawdataDF<-rawdataDF[which(rawdataDF$DecisionMinor!="recommend after revision"),]
rawdataDF<-rawdataDF[which(rawdataDF$DecisionMinor!="recommended after revision"),]
rawdataDF<-rawdataDF[which(rawdataDF$DecisionMinor!="revisions required"),]


rawdataDF$DecisionMinor <- factor(rawdataDF$DecisionMinor, levels = c("accept","minor revision","major revision","reject"))

inputDF <- rawdataDF
inputDF<-inputDF[which(inputDF$CorrespondingAuthorGender!="U"),]
inputDF<-inputDF[which(inputDF$ReviewerGender!="U"),]
inputDF$CorrespondingAuthorGender <- factor(inputDF$CorrespondingAuthorGender, levels = c("F","M"))
inputDF$ReviewerGender <- factor(inputDF$ReviewerGender, levels = c("F","M"))

modelToUse3 <- multinom(DecisionMinor ~ ReviewerGender*CorrespondingAuthorGender, data = inputDF, trace=FALSE)

# need to also fit the model with one variable only for comparison for the chi square test (anova only works on multiple multinom output objects)
modelToUse2 <- multinom(DecisionMinor ~ ReviewerGender, data = inputDF, trace=FALSE)

anovasummary<-anova(modelToUse3,modelToUse2,test="Chisq")
PValueForModel<-anovasummary[2,"Pr(Chi)"]
SignificanceForModel<-GetSignificanceFromPValue(PValueForModel)

effect("ReviewerGender*CorrespondingAuthorGender",data=inputDF,mod=modelToUse3,confidence.level=.95)->rawEffects

resultsDF<- addPopulationSizeToDataFrame(rawEffects,modelToUse3,"ReviewerGender","CorrespondingAuthorGender")

t1<- melt(resultsDF[,c("ReviewerGender","CorrespondingAuthorGender", "prob.accept","prob.minor.revision", "prob.major.revision","prob.reject")], id=c("ReviewerGender","CorrespondingAuthorGender"))
colnames(t1)<-c("ReviewerGender","CorrespondingAuthorGender","Effect","fit")
levels(t1$Effect) <- c("accept","minor.revision","major.revision","reject")
t2<- melt(resultsDF[,c("ReviewerGender","CorrespondingAuthorGender", "L.prob.accept","L.prob.minor.revision", "L.prob.major.revision","L.prob.reject")], id=c("ReviewerGender","CorrespondingAuthorGender"))
colnames(t2)<-c("ReviewerGender","CorrespondingAuthorGender","Effect","Lower")
levels(t2$Effect) <- c("accept","minor.revision","major.revision","reject")
plotableTable<- merge(t1, t2, by = c("ReviewerGender", "CorrespondingAuthorGender","Effect"),all =TRUE)

t3<- melt(resultsDF[,c("ReviewerGender","CorrespondingAuthorGender", "U.prob.accept","U.prob.minor.revision", "U.prob.major.revision","U.prob.reject")], id=c("ReviewerGender","CorrespondingAuthorGender"))
colnames(t3)<-c("ReviewerGender","CorrespondingAuthorGender","Effect","Upper")
levels(t3$Effect) <- c("accept","minor.revision","major.revision","reject")
plotableTable<- merge(plotableTable, t3, by = c("ReviewerGender", "CorrespondingAuthorGender","Effect"),all =TRUE)

t4<- melt(resultsDF[,c("ReviewerGender","CorrespondingAuthorGender", "se.prob.accept","se.prob.minor.revision", "se.prob.major.revision","se.prob.reject")], id=c("ReviewerGender","CorrespondingAuthorGender"))
colnames(t4)<-c("ReviewerGender","CorrespondingAuthorGender","Effect","SE")
levels(t4$Effect) <- c("accept","minor.revision","major.revision","reject")
plotableTable<- merge(plotableTable, t4, by = c("ReviewerGender", "CorrespondingAuthorGender","Effect"),all =TRUE)

ggplot(plotableTable,aes(y=fit,ymin=Lower,ymax=Upper,x=CorrespondingAuthorGender,group=1))+
  geom_point()+
  geom_line()+
  geom_errorbar(alpha=0.7,width=0.1)+
  theme(legend.position="bottom",legend.direction="horizontal")+
  ylab("Proportion of reviewer recommendations")+
  xlab("Gender of corresponding author")+
  ggtitle(paste0("Breakdown by Reviewer gender in separate row plots"))+
  facet_grid(vars(ReviewerGender), vars(Effect))+theme_bw()

```

Figure - Multinomial GLM model of reviewer recommendations, corresponding author gender and reviewer gender (female reviewers in top plots and male reviewers in lower plots) (Model: ReviewerRecommendation ~ ReviewerGender \* CorrAuthorGender). ANOVA: ReviewerGender \* CorrespondingAuthorGender is  `r toString(SignificanceForModel)` (p = `r toString(PValueForModel)`)

```{r echo=FALSE,warning=FALSE,message=FALSE}

kable(plotableTable)
